\scshape Advanced game theory

### 2025-12-07 ch_aps.tex
Add files via upload

\begin{definition}
	%
	\label{definition:stage_game}
	%
	A \emph{public-monitoring stage game} is a tuple $\left( I, Y, (A_i, u_i)_{i \in I}, \mu \right)$, where $I$ and $Y$ are non-empty sets, for each $i \in I$, $A_i$ is a non-empty set and $u_i$ is a function $A_i \times Y \to \R$, and $\mu$ is a function $A \to \Delta(Y)$.
	%
\end{definition}

@@

\begin{remark}
	%
	\label{remark:rep_game_formal}
	%
	As mentioned above, for a finite public-information stage game $G = \left( I, Y, (A_i, u_i)_{i \in I}, \mu \right)$ and a $\delta \in (0,1)$, the public-monitoring repeated game $(G,\delta)$ is formally an extensive-form game, namely
	%
	\begin{equation*}
		\left(\widetilde{I},\widetilde{\mathcal{A}},\widetilde{H},\widetilde{P},\widetilde{A},\widetilde{\mathcal{S}},\left(\widetilde{u}_i\right)_{i \in I},\widetilde{\pi}\right)
	\end{equation*}
	%
	where
	%
	\begin{itemize}

@@

		\item the player and action functions $\widetilde{P}$ and $\widetilde{A}$ are given by

@@

			\item $\widetilde{P}(h) \coloneqq 1$ and $\widetilde{A}(h) \coloneqq A_1$ if $\text{length}(h) = (\abs*{I}+1) n$ for some $n \in \{0,1,2,3,\dots\}$,

			\item $\widetilde{P}(h) \coloneqq 2$ and $\widetilde{A}(h) \coloneqq A_2$ if $\text{length}(h) = (\abs*{I}+1) n + 1$ for some $n \in \{0,1,2,3,\dots\}$,

@@

			\item $\widetilde{P}(h) \coloneqq \abs*{I}$ and $\widetilde{A}(h) \coloneqq A_{\abs*{I}}$ if $\text{length}(h) = (\abs*{I}+1) n + \abs*{I}-1$ for some $n \in \{0,1,2,3,\dots\}$, and

			\item $\widetilde{P}(h) \coloneqq \text{Nature}$ and $\widetilde{A}(h) \coloneqq Y$ if $\text{length}(h) = (\abs*{I}+1) n + \abs*{I}$ for some $n \in \{0,1,2,3,\dots\}$,

@@

A \emph{(behavioural) strategy} of player $i \in I$ is a map $\sigma_i : \Union_{t=0}^\infty (Y \times A_i)^t \to \Delta(A_i)$ which specifies a (mixed) action at each private history of player~$i$. A (behavioural) strategy $\sigma_i$ of player $i \in I$ is called \emph{public} iff it is measurable with respect to the public history, i.e. there exists a function $\tau_i : \Union_{t=0}^\infty Y^t \to \Delta(A_i)$ such that
%
\begin{equation*}
	\sigma_i\left(y^0,a^0_i,y^1,a^1_i,y^2,a^2_i,\dots,y^{t-1},a^{t-1}_i\right)
	= \tau_i\left( y^0, y^1, y^2, \dots, y^{t-1} \right)
\end{equation*}
%
for every private history
%
\begin{equation*}
	\left(y^0,a^0_i,y^1,a^1_i,y^2,a^2_i,\dots,y^{t-1},a^{t-1}_i\right)
	\in (Y \times A_i)^t .
\end{equation*}
%
A \emph{(public) strategy profile} is a collection $\sigma = (\sigma_i)_{i \in I}$ of (public) strategies, one for each player. For a public strategy profile $\sigma$, we abuse notation by writing $\sigma\bigl(h^{t-1}\bigr) \in \prod_{i \in I} \Delta(A_i)$ for the mixed action profile played at public history $h^{t-1} \in Y^t$. (Why is this an abuse of notation?)

Any public strategy profile $\sigma$ generates, for each $t \in \{0,1,2,3,\dots\}$, a random length-$t$ public history $h^{t-1}_\sigma \in Y^t$: in particular, $h^{-1}_\sigma \coloneqq \varnothing$ (the empty sequence), $h^0_\sigma \coloneqq (y^0_\sigma)$ where $y^0_\sigma$ is drawn from $\mu\left( \cdot \middle| \sigma\left( h^{-1}_\sigma \right) \right)$, $h^1_\sigma \coloneqq \left(h^0_\sigma,y^1_\sigma\right)$ where $y^1_\sigma$ is drawn from $\mu\left( \cdot \middle| \sigma\left( h^0_\sigma \right) \right)$, $h^2_\sigma \coloneqq \left(h^1_\sigma,y^2_\sigma\right)$ where $y^2_\sigma$ is drawn from $\mu\left( \cdot \middle| \sigma\left( h^1_\sigma \right) \right)$, and so on, where the draws are jointly independent. The (lifetime) expected payoff of player $i \in I$ from the strategy profile $\sigma$ is denoted
%
\begin{equation*}
	U_i(\sigma) \coloneqq (1-\delta) \sum_{t=0}^\infty \delta^t u_i\left( \sigma\left(h^{t-1}_\sigma\right) \right) .
\end{equation*}
%
We write $U(\sigma) \coloneqq (U_i(\sigma))_{i \in I}$ for the vector of all players' lifetime payoffs. Similarly, we write $u(\alpha) \coloneqq (u_i(\alpha))_{i \in I}$ for the vector of all players' stage-game payoffs from a mixed action profile $\alpha \in \prod_{i \in I} \Delta(A_i)$.

@@

The \emph{continuation payoff} of a public strategy profile $\sigma$ at a public history $h_{*}^{t-1} \in Y^t$ is the payoff from play of the (public) continuation strategy profile $\sigma_{*} \coloneqq \sigma|_{h_{*}^{t-1}}$ in the `continuation game' that starts at $h_{*}^{t-1}$, i.e.
%
\begin{align*}
	U\left( \sigma_{*} \right)
	&= (1-\delta) \sum_{s=0}^\infty \delta^s u\left( \sigma_{*}\left(h^{s-1}_{\sigma_{*}}\right) \right)
	\\
	&= (1-\delta) \sum_{s=t}^\infty \delta^{s-t} u\left( \sigma\left(h_{*}^{t-1},h^{s-t-1}_{\sigma_{*}} \right) \right) .
\end{align*}
%
Note that by definition, the continuation payoff from public history $h_{*}^{t-1} \in Y^t$ omits per-period payoffs enjoyed in periods $\{0,1,2,3,\dots,t-1\}$; it is a forward-looking object.

@@

\begin{corollary}
	%
	\label{corollary:enforcement}
	%
	Fix a map $\beta : Y \to E$ and a mixed action profile $\alpha \in \prod_{i \in I} \Delta(A_i)$, and let
	%
	\begin{equation*}
		v \coloneqq (1-\delta) u(\alpha) + \delta \sum_{y \in Y} \beta(y) \mu(y|\alpha) .
	\end{equation*}
	%
	If
	%
	\begin{equation}
		v_i \geq (1-\delta) u_i(a_i,\alpha_{-i}) + \delta \sum_{y \in Y} \beta_i(y) \mu(y|a_i,\alpha_{-i})
		\label{eq:enforcement_cor}
	\end{equation}
	%
	for each player $i \in I$ and action $a_i \in A_i$, then $v \in E$.
	%
\end{corollary}

@@

\begin{definition}
	%
	\label{definition:enforcement}
	%
	A map $\beta : Y \to X$ \emph{enforces} a mixed action profile $\alpha \in \prod_{i \in I} \Delta(A_i)$ iff
	%
	\begin{align*}
		(1-\delta) u_i(\alpha) &+ \delta \sum_{y \in Y} \beta_i(y) \mu(y|\alpha)
		\\
		\geq (1-\delta) u_i(a_i,\alpha_{-i}) &+ \delta \sum_{y \in Y} \beta_i(y) \mu(y|a_i,\alpha_{-i})
	\end{align*}
	%
	for each player $i \in I$ and action $a_i \in A_i$.
	%
\end{definition}

@@

\begin{definition}
	%
	\label{definition:generation}
	%
	A set $V \subseteq X$ of payoff vectors \emph{generates} a payoff vector $v \in X$ iff there exists a map $\beta : Y \to V$ and a mixed action profile $\alpha \in \prod_{i \in I} \Delta(A_i)$ such that $\beta$ enforces $\alpha$ and
	%
	\begin{equation*}
		(1-\delta) u(\alpha)
		+ \delta \sum_{y \in Y} \beta(y) \mu(y|\alpha) 
		= v.
	\end{equation*}
	%
\end{definition}

@@

		\item For the length-$0$ public history $\varnothing$, since $v \in V \subseteq B(V)$, there exists a map $\beta_{\varnothing} : Y \to V$ and a mixed action profile $\alpha_{\varnothing} \in \prod_{i \in I} \Delta(A_i)$ such that $\beta_{\varnothing}$ enforces $\alpha_{\varnothing}$ and
		%
		\begin{equation*}
			(1-\delta) u\left(\alpha_{\varnothing}\right)
			+ \delta \sum_{y \in Y} \beta_{\varnothing}(y) \mu\left(y\middle|\alpha_{\varnothing}\right) 
			= v ;
		\end{equation*}
		%
		let $\sigma\left( \varnothing \right) \coloneqq \alpha_{\varnothing}$.

		\item For each length-$1$ public history $h^0 \equiv \left(y^0\right) \in Y$, since $\beta_{\varnothing}\left( y^0 \right) \in V \subseteq B(V)$, there exists a map $\beta_{h^0} : Y \to V$ and a mixed action profile $\alpha_{h^0} \in \prod_{i \in I} \Delta(A_i)$ such that $\beta_{h^0}$ enforces $\alpha_{h^0}$ and
		%
		\begin{equation*}
			(1-\delta) u\left(\alpha_{h^0}\right)
			+ \delta \sum_{y \in Y} \beta_{h^0}(y) \mu\left(y\middle|\alpha_{h^0}\right) 
			= \beta_{\varnothing}\left( y^0 \right) ;
		\end{equation*}
		%
		let $\sigma\left( h^0 \right) \coloneqq \alpha_{h^0}$.

		\item For each length-$2$ public history $h^1 \equiv \left(h^0,y^1\right) \in Y^2$, since $\beta_{h^0}\left( y^1 \right) \in V \subseteq B(V)$, there exists a map $\beta_{h^1} : Y \to V$ and a mixed action profile $\alpha_{h^1} \in \prod_{i \in I} \Delta(A_i)$ such that $\beta_{h^1}$ enforces $\alpha_{h^1}$ and
		%
		\begin{equation*}
			(1-\delta) u\left(\alpha_{h^1}\right)
			+ \delta \sum_{y \in Y} \beta_{h^1}(y) \mu\left(y\middle|\alpha_{h^1}\right) 
			= \beta_{h^0}\left( y^1 \right) ;
		\end{equation*}
		%
		let $\sigma\left( h^1 \right) \coloneqq \alpha_{h^1}$.

		\item For each length-$3$ public history $h^2 \equiv \left(h^1,y^2\right) \in Y^3$, since $\beta_{h^1}\left( y^2 \right) \in V \subseteq B(V)$, there exists a map $\beta_{h^2} : Y \to V$ and a mixed action profile $\alpha_{h^2} \in \prod_{i \in I} \Delta(A_i)$ such that $\beta_{h^2}$ enforces $\alpha_{h^2}$ and
		%
		\begin{equation*}
			(1-\delta) u\left(\alpha_{h^2}\right)
			+ \delta \sum_{y \in Y} \beta_{h^2}(y) \mu\left(y\middle|\alpha_{h^2}\right) 
			= \beta_{h^1}\left( y^2 \right) ;
		\end{equation*}
		%
		let $\sigma\left( h^2 \right) \coloneqq \alpha_{h^2}$.

@@

\begin{remark}
	%
	\label{remark:aps_randomisation}
	%
	Recall from \cref{aps:feasible} (\cpageref{item:X_converse:stage_pos}) the notion of a \emph{public correlating device.} Throughout this section, we have (implicitly) assumed that players do \emph{not} have access to a public correlating device, by always considering only mixed action profiles $\alpha \in \prod_{i \in I} \Delta(A_i)$ in which players mix \emph{independently} (as opposed to considering all mixed action profiles $\alpha \in \Delta(A)$, which may feature correlation between players' actions). The (only) place where this matters is in how `enforcement' is defined. In particular, when there \emph{is} a public correlating device, we say that $\beta : Y \to X$ \emph{enforces} a mixed action profile $\alpha \in \Delta(A)$ (in which mixing may be correlated across players) iff $\alpha$ is a \emph{correlated} equilibrium of the (artificial) normal-form game $\left(I,(A_i,w_i\right)_{i \in I})$ from \Cref{remark:enforc_nash} (\cpageref{remark:enforc_nash} above). With enforcement defined in this way, the three results of this section (\Cref{proposition:self-gen}, \Cref{theorem:self-gen_max_ppe} and \Cref{proposition:ppe_algo}) remain true exactly as stated.%
		\footnote{An alternative way of adding a public correlating device, which you'll find in chapter~7 of \textcite{MailathSamuelson2006}, is to encode it in the monitoring structure, by expanding the set of public outcomes to $Y' \coloneqq Y \times [0,1]$ and letting $\mu' \coloneqq \mu \times \lambda$, where $\lambda$ is the Lebesgue measure; in other words, public outcomes are now $(y,z) \in Y \times [0,1]$, where $z$ is a uniformly distributed random variable that is independent of $y$. This way of modelling a public correlating device does \emph{not} require the definition of enforcement to be altered. But it seems to me that this approach is not quite equivalent, because it does not allow actions to be correlated in period $t=0$.}
	That doesn't mean that it makes no difference whether or not there is a public correlating device: see the next exercise, and the final paragraph of the next section.
	%
\end{remark}

@@

	\item \label{item:does_max1} For a given player $i \in I$, what are the lowest and highest payoff of player~$i$ in a PPE? That is, what are the values of $\min_{v \in E} v_i$ and $\max_{v \in E} v_i$?

@@

\begin{exercise}[S.~Petersen]
	%
	\label{exercise:ir_ppe_necessary_pure}
	%
	Prove by example that a PPE payoff vector $v \in E$ need \emph{not} be individually rational in the stronger sense that $v_i \geq \min_{a_{-i} \in A_{-i}} \max_{a_i \in A_i} u_i(a_i,a_{-i})$ for every player $i \in I$.
	%
\end{exercise}

>>>>

\begin{definition}
	%
	\label{definition:stage_game}
	%
	A \emph{public-monitoring stage game} is a tuple
	%
	\begin{equation*}
		\left( I, Y, (A_i, u_i)_{i \in I}, \mu \right) ,
	\end{equation*}
	%
	where $I$ and $Y$ are non-empty sets, for each $i \in I$, $A_i$ is a non-empty set and $u_i$ is a function $A_i \times Y \to \R$, and $\mu$ is a function $A \to \Delta(Y)$.
	%
\end{definition}

@@

\begin{remark}
	%
	\label{remark:rep_game_formal}
	%
	As mentioned above, for a finite public-information stage game $G = \left( I, Y, (A_i, u_i)_{i \in I}, \mu \right)$ and a $\delta \in (0,1)$, the public-monitoring repeated game $(G,\delta)$ is formally an extensive-form game, namely
	%
	\begin{equation*}
		\left(\widetilde{I},\widetilde{\mathcal{A}},\widetilde{H},\widetilde{P},\widetilde{\mathcal{S}},\left(\widetilde{u}_i\right)_{i \in I},\widetilde{\pi}\right)
	\end{equation*}
	%
	where
	%
	\begin{itemize}

@@

		\item the action and player functions $\widetilde{A}$ and $\widetilde{P}$ are given by

@@

			\item $\widetilde{A}(h) \coloneqq A_1$ and $\widetilde{P}(h) \coloneqq 1$ if $\text{length}(h) = (\abs*{I}+1) n$ for some $n \in \{0,1,2,3,\dots\}$,

			\item $\widetilde{A}(h) \coloneqq A_2$ and $\widetilde{P}(h) \coloneqq 2$ if $\text{length}(h) = (\abs*{I}+1) n + 1$ for some $n \in \{0,1,2,3,\dots\}$,

@@

			\item $\widetilde{A}(h) \coloneqq A_{\abs*{I}}$ and $\widetilde{P}(h) \coloneqq \abs*{I}$ if $\text{length}(h) = (\abs*{I}+1) n + \abs*{I}-1$ for some $n \in \{0,1,2,3,\dots\}$, and

			\item $\widetilde{A}(h) \coloneqq Y$ and $\widetilde{P}(h) \coloneqq \text{Nature}$ if $\text{length}(h) = (\abs*{I}+1) n + \abs*{I}$ for some $n \in \{0,1,2,3,\dots\}$,

@@

A \emph{(behavioural) strategy} of player $i \in I$ is a map $\sigma_i : \Union_{t=0}^\infty (Y \times A_i)^t \to \Delta(A_i)$ which specifies a (mixed) action at each private history of player~$i$. A (behavioural) strategy $\sigma_i$ of player $i \in I$ is called \emph{public} iff it is measurable with respect to the public history, i.e. there exists a function $\tau_i : \Union_{t=0}^\infty Y^t \to \Delta(A_i)$ such that
%
\begin{equation*}
	\sigma_i\left(y^0,a^0_i,y^1,a^1_i,y^2,a^2_i,\dots,y^{t-1},a^{t-1}_i\right)
	= \tau_i\left( y^0, y^1, y^2, \dots, y^{t-1} \right)
\end{equation*}
%
for every private history
%
\begin{equation*}
	\left(y^0,a^0_i,y^1,a^1_i,y^2,a^2_i,\dots,y^{t-1},a^{t-1}_i\right)
	\in (Y \times A_i)^t .
\end{equation*}
%
A \emph{(public) strategy profile} is a collection $\sigma = (\sigma_i)_{i \in I}$ of (public) strategies, one for each player. For a public strategy profile $\sigma$, we abuse notation by writing $\sigma\bigl(h^{t-1}\bigr) \in \prod_{i \in I} \Delta(A_i)$ for the profile of mixed actions played at public history $h^{t-1} \in Y^t$. (Why is this an abuse of notation?)

Any public strategy profile $\sigma$ generates, for each $t \in \{0,1,2,3,\dots\}$, a random length-$t$ public history $h^{t-1}_\sigma \in Y^t$: in particular, $h^{-1}_\sigma \coloneqq \varnothing$ (the empty sequence), $h^0_\sigma \coloneqq (y^0_\sigma)$ where $y^0_\sigma$ is drawn from $\mu\left( \cdot \middle| \sigma\left( h^{-1}_\sigma \right) \right)$, $h^1_\sigma \coloneqq \left(h^0_\sigma,y^1_\sigma\right)$ where $y^1_\sigma$ is drawn from $\mu\left( \cdot \middle| \sigma\left( h^0_\sigma \right) \right)$, $h^2_\sigma \coloneqq \left(h^1_\sigma,y^2_\sigma\right)$ where $y^2_\sigma$ is drawn from $\mu\left( \cdot \middle| \sigma\left( h^1_\sigma \right) \right)$, and so on, where the draws are jointly independent. The lifetime expected payoff of player $i \in I$ from the strategy profile $\sigma$ is denoted
%
\begin{equation*}
	U_i(\sigma) \coloneqq \E\left( (1-\delta) \sum_{t=0}^\infty \delta^t u_i\left( \sigma\left(h^{t-1}_\sigma\right) \right) \right) .
\end{equation*}
%
We write $U(\sigma) \coloneqq (U_i(\sigma))_{i \in I}$ for the vector of all players' lifetime expected payoffs. Similarly, we write $u(\alpha) \coloneqq (u_i(\alpha))_{i \in I}$ for the vector of all players' stage-game payoffs from a profile $\alpha \in \prod_{i \in I} \Delta(A_i)$ of mixed actions.

@@

The \emph{continuation payoff} of a public strategy profile $\sigma$ at a public history $h_{*}^{t-1} \in Y^t$ is the payoff from play of the (public) continuation strategy profile $\sigma_{*} \coloneqq \sigma|_{h_{*}^{t-1}}$ in the `continuation game' that starts at $h_{*}^{t-1}$, i.e.
%
\begin{align*}
	U\left( \sigma_{*} \right)
	&= \E\left( (1-\delta) \sum_{s=0}^\infty \delta^s u\left( \sigma_{*}\left(h^{s-1}_{\sigma_{*}}\right) \right) \right)
	\\
	&= \E\left( (1-\delta) \sum_{s=t}^\infty \delta^{s-t} u\left( \sigma\left(h_{*}^{t-1},h^{s-t-1}_{\sigma_{*}} \right) \right) \right) .
\end{align*}
%
Note that by definition, the continuation payoff from public history $h_{*}^{t-1} \in Y^t$ omits per-period payoffs enjoyed in periods $\{0,1,2,3,\dots,t-1\}$; it is a forward-looking object.

@@

\begin{corollary}
	%
	\label{corollary:enforcement}
	%
	Fix a map $\beta : Y \to E$ and a profile $\alpha \in \prod_{i \in I} \Delta(A_i)$ of mixed actions, and let
	%
	\begin{equation*}
		v \coloneqq (1-\delta) u(\alpha) + \delta \sum_{y \in Y} \beta(y) \mu(y|\alpha) .
	\end{equation*}
	%
	If
	%
	\begin{equation}
		v_i \geq (1-\delta) u_i(a_i,\alpha_{-i}) + \delta \sum_{y \in Y} \beta_i(y) \mu(y|a_i,\alpha_{-i})
		\label{eq:enforcement_cor}
	\end{equation}
	%
	for each player $i \in I$ and action $a_i \in A_i$, then $v \in E$.
	%
\end{corollary}

@@

\begin{definition}
	%
	\label{definition:enforcement}
	%
	A map $\beta : Y \to X$ \emph{enforces} a profile $\alpha \in \prod_{i \in I} \Delta(A_i)$ of mixed actions iff
	%
	\begin{align*}
		(1-\delta) u_i(\alpha) &+ \delta \sum_{y \in Y} \beta_i(y) \mu(y|\alpha)
		\\
		\geq (1-\delta) u_i(a_i,\alpha_{-i}) &+ \delta \sum_{y \in Y} \beta_i(y) \mu(y|a_i,\alpha_{-i})
	\end{align*}
	%
	for each player $i \in I$ and action $a_i \in A_i$.
	%
\end{definition}

@@

\begin{definition}
	%
	\label{definition:generation}
	%
	A set $V \subseteq X$ of payoff vectors \emph{generates} a payoff vector $v \in X$ iff there exists a map $\beta : Y \to V$ and a profile $\alpha \in \prod_{i \in I} \Delta(A_i)$ of mixed actions such that $\beta$ enforces $\alpha$ and
	%
	\begin{equation*}
		(1-\delta) u(\alpha)
		+ \delta \sum_{y \in Y} \beta(y) \mu(y|\alpha) 
		= v.
	\end{equation*}
	%
\end{definition}

@@

		\item For the length-$0$ public history $\varnothing$, since $v \in V \subseteq B(V)$, there exists a map $\beta_{\varnothing} : Y \to V$ and a profile $\alpha_{\varnothing} \in \prod_{i \in I} \Delta(A_i)$ of mixed actions such that $\beta_{\varnothing}$ enforces $\alpha_{\varnothing}$ and
		%
		\begin{equation*}
			(1-\delta) u\left(\alpha_{\varnothing}\right)
			+ \delta \sum_{y \in Y} \beta_{\varnothing}(y) \mu\left(y\middle|\alpha_{\varnothing}\right) 
			= v ;
		\end{equation*}
		%
		let $\sigma\left( \varnothing \right) \coloneqq \alpha_{\varnothing}$.

		\item For each length-$1$ public history $h^0 \equiv \left(y^0\right) \in Y$, since $\beta_{\varnothing}\left( y^0 \right) \in V \subseteq B(V)$, there exists a map $\beta_{h^0} : Y \to V$ and a profile $\alpha_{h^0} \in \prod_{i \in I} \Delta(A_i)$ of mixed actions such that $\beta_{h^0}$ enforces $\alpha_{h^0}$ and
		%
		\begin{equation*}
			(1-\delta) u\left(\alpha_{h^0}\right)
			+ \delta \sum_{y \in Y} \beta_{h^0}(y) \mu\left(y\middle|\alpha_{h^0}\right) 
			= \beta_{\varnothing}\left( y^0 \right) ;
		\end{equation*}
		%
		let $\sigma\left( h^0 \right) \coloneqq \alpha_{h^0}$.

		\item For each length-$2$ public history $h^1 \equiv \left(h^0,y^1\right) \in Y^2$, since $\beta_{h^0}\left( y^1 \right) \in V \subseteq B(V)$, there exists a map $\beta_{h^1} : Y \to V$ and a profile $\alpha_{h^1} \in \prod_{i \in I} \Delta(A_i)$ of mixed actions such that $\beta_{h^1}$ enforces $\alpha_{h^1}$ and
		%
		\begin{equation*}
			(1-\delta) u\left(\alpha_{h^1}\right)
			+ \delta \sum_{y \in Y} \beta_{h^1}(y) \mu\left(y\middle|\alpha_{h^1}\right) 
			= \beta_{h^0}\left( y^1 \right) ;
		\end{equation*}
		%
		let $\sigma\left( h^1 \right) \coloneqq \alpha_{h^1}$.

		\item For each length-$3$ public history $h^2 \equiv \left(h^1,y^2\right) \in Y^3$, since $\beta_{h^1}\left( y^2 \right) \in V \subseteq B(V)$, there exists a map $\beta_{h^2} : Y \to V$ and a profile $\alpha_{h^2} \in \prod_{i \in I} \Delta(A_i)$ of mixed actions such that $\beta_{h^2}$ enforces $\alpha_{h^2}$ and
		%
		\begin{equation*}
			(1-\delta) u\left(\alpha_{h^2}\right)
			+ \delta \sum_{y \in Y} \beta_{h^2}(y) \mu\left(y\middle|\alpha_{h^2}\right) 
			= \beta_{h^1}\left( y^2 \right) ;
		\end{equation*}
		%
		let $\sigma\left( h^2 \right) \coloneqq \alpha_{h^2}$.

@@

\begin{remark}
	%
	\label{remark:aps_randomisation}
	%
	Recall from \cref{aps:feasible} (\cpageref{item:X_converse:stage_pos}) the notion of a \emph{public correlating device.} Throughout this section, we have (implicitly) assumed that players do \emph{not} have access to a public correlating device, by always considering only action profiles $\alpha \in \prod_{i \in I} \Delta(A_i)$ in which players mix \emph{independently} (as opposed to considering all mixed action profiles $\alpha \in \Delta(A)$, which may feature correlation between players' actions). The (only) place where this matters is in how `enforcement' is defined. In particular, when there \emph{is} a public correlating device, we say that $\beta : Y \to X$ \emph{enforces} a mixed action profile $\alpha \in \Delta(A)$ (in which mixing may be correlated across players) iff $\alpha$ is a \emph{correlated} equilibrium of the (artificial) normal-form game $\left(I,(A_i,w_i\right)_{i \in I})$ from \Cref{remark:enforc_nash} (\cpageref{remark:enforc_nash} above). With enforcement defined in this way, the three results of this section (\Cref{proposition:self-gen}, \Cref{theorem:self-gen_max_ppe} and \Cref{proposition:ppe_algo}) remain true exactly as stated.%
		\footnote{An alternative way of adding a public correlating device, which you'll find in chapter~7 of \textcite{MailathSamuelson2006}, is to encode it in the monitoring structure, by expanding the set of public outcomes to $Y' \coloneqq Y \times [0,1]$ and letting $\mu' \coloneqq \mu \times \lambda$, where $\lambda$ is the Lebesgue measure; in other words, public outcomes are now $(y,z) \in Y \times [0,1]$, where $z$ is a uniformly distributed random variable that is independent of $y$. This way of modelling a public correlating device does \emph{not} require the definition of enforcement to be altered. But it seems to me that this approach is not quite equivalent, because it does not allow actions to be correlated in period $t=0$.}
	That doesn't mean that it makes no difference whether or not there is a public correlating device: see the next exercise, and the final paragraph of the next section.
	%
\end{remark}

@@

	\item \label{item:does_max1} For a given player $i \in I$, what is the lowest and highest payoff of player~$i$ in a PPE? That is, what are the values of $\min_{v \in E} v_i$ and $\max_{v \in E} v_i$?

@@

\begin{exercise}[Sami Petersen]
	%
	\label{exercise:ir_ppe_necessary_pure}
	%
	Prove by example that a PPE payoff vector $v \in E$ need \emph{not} be individually rational in the stronger sense that $v_i \geq \min_{a_{-i} \in A_{-i}} \max_{a_i \in A_i} u_i(a_i,a_{-i})$ for every player $i \in I$.
	%
\end{exercise}

### 2025-12-07 ch_dom.tex
Add files via upload

Many (most?) games arising in economics are not dominance-solvable.

@@

\hyperref[lemma:pearce]{Pearce's lemma} is from \textcite[][Lemma~3]{Pearce1984}. Similar results were known earlier; according to Bruno \textcite{Salcedo2020}, the earliest are those of \textcite{Wald1939,Stein1955}. \hyperref[lemma:pearce]{Pearce's lemma} is an instance of a more general (often useful) principle that `undominated' is equivalent to `optimal in some situation' \parencite[see e.g.][]{BorgersCheng2023}. The proof in the text (based on the separating-hyperplane theorem) is standard, but does not appear in \textcite{Pearce1984}; \textcite{Salcedo2020} attributes it to Abraham Wald.

@@

>>>>

In many (most?) games that arise in economics, rationalisability does not yield a unique prediction of play.

@@

\hyperref[lemma:pearce]{Pearce's lemma} is from \textcite[][Lemma~3]{Pearce1984}. Similar results were known earlier; according to Bruno \textcite{Salcedo2020}, the earliest are those of \textcite{Wald1939,Stein1955}. \hyperref[lemma:pearce]{Pearce's lemma} is an instance of a more general (often useful) principle that `undominated' is equivalent to `optimal in some situation' \parencite[see e.g.][]{Fishburn1975,BorgersCheng2023}. The proof in the text (based on the separating-hyperplane theorem) is standard, but does not appear in \textcite{Pearce1984}; \textcite{Salcedo2020} attributes it to Abraham Wald.

@@

\begin{exercise}
	%
	\label{exercise:exam_ht25}
	%
	Consider the Bertrand model of oligopolistic competition, except with constraints on what prices can be charged. Two firms $i \in I \coloneqq \{1,2\}$ sell a homogeneous good, and choose what price $a_i \in A_i$ to charge, where $A_i$ is a non-empty subset of $[0,\infty)$. There is a single consumer, who buys from whichever firm charges less; if prices are equal, then the consumer randomises by tossing a fair coin. Firms' marginal costs are equal and constant, normalised to zero. Formally, the profit of each firm $i \in \{1,2\}$ is
	%
	\begin{equation*}
		u_i(a_i,a_{3-i})
		\coloneqq
		\begin{cases}
			a_i	& \text{if $a_i < a_{3-i}$}
			\\
			a_i/2	& \text{if $a_i = a_{3-i}$}
			\\
			0	& \text{if $a_i > a_{3-i}$.}
		\end{cases}
	\end{equation*}
	%
	Some reminders, in case you need them:

	\begin{itemize}

		\item `$\Delta(A_i)$' denotes the set of all probability distributions on $A_i$.

		\item An action $a_i \in A_i$ is a \emph{best reply} to a mixed action $\alpha_{3-i} \in \Delta(A_{3-i})$ if and only if $u_i(a_i,\alpha_{3-i}) \geq u_i(a_i',\alpha_{3-i})$ for every pure action $a_i' \in A_i$, where
		%
		\begin{equation*}
			u_i(a_i,\alpha_{3-i})
			\equiv \sum_{a_{3-i} \in A_{3-i}} u_i(a_i,a_{3-i}) \alpha_{3-i}(a_{3-i}) .
		\end{equation*}

		\item An action $a_i \in A_i$ of a player $i \in \{1,2\}$ is called \emph{never-best} if and only if there is no mixed action $\alpha_{3-i} \in \Delta(A_{3-i})$ to which $a_i$ is a best reply.

		\item An action $a_i \in A_i$ of a player $i \in \{1,2\}$ is \emph{strictly dominated} by a mixed action $\alpha_i \in \Delta(A_i)$ if and only if $u_i(a_i,a_{3-i}) < u_i(\alpha_i,a_{3-i})$ for every pure action $a_{3-i} \in A_{3-i}$, where
		%
		\begin{equation*}
			u_i(\alpha_i,a_{3-i})
			\equiv \sum_{a_i \in A_i} u_i(a_i,a_{3-i}) \alpha_i(a_i) .
		\end{equation*}

		\item Pearce's lemma says that if the action sets $A_1$ and $A_2$ are finite, then an action $a_i \in A_i$ of a player $i \in \{1,2\}$ is never-best if and only if it is strictly dominated by some mixed action $\alpha_i \in \Delta(A_i)$.

		\item The \emph{rationalisable} actions $X_1^\infty \subseteq A_1$ and $X_2^\infty \subseteq A_2$ are those actions that are left after all actions of either player that are never-best are deleted, and then all actions of either player that are (now) never-best are deleted, and so on forever.

	\end{itemize}
	%
	Answer the following.

		\item Suppose that $0 \in A_1$ and $0 \in A_2$.

			\item Show that no action is never-best. (That is, there is no player $i \in \{1,2\}$ and action $a_i \in A_i$ such that $a_i$ is never-best.)

			\item What are the rationalisable actions $X_1^\infty$ and $X_2^\infty$?

		\item Suppose that $A_1 = A_2 = \{1,2,3,4,\dots,K-1,K\}$, where $K$ is an integer that satisfies $K \geq 3$. (Note that $0 \notin A_1$ and $0 \notin A_2$.)

			\item Show that $(a_1,a_2) = (2,2)$ is a Nash equilibrium.

			\item Show that $a_i = K$ is \emph{not} strictly dominated by the pure action $K-1$.

			\item Show that $a_i = K$ is never-best. (Take care.)

			\item What are the rationalisable actions $X_1^\infty$ and $X_2^\infty$? (Take care.)

		\item Suppose that $A_1 = A_2 = [k,\infty)$ where $k>0$. Let $F_k$ denote the CDF $F_k : [0,\infty) \to [0,1]$ defined by
		%
		\begin{equation*}
			F_k(x) \coloneqq
			\begin{cases}
				0 & \text{for $x \in [0,k)$} \\
				1 - k/x & \text{for $x \in [k,\infty)$.}
			\end{cases}
		\end{equation*}
		%
		($F_k$ is called the `$\text{Pareto}(k,1)$ distribution'.)

			\item Show that $(F_k,F_k)$ is a Nash equilibrium.

			\item What are the rationalisable actions $X_1^\infty$ and $X_2^\infty$?
